Supplementary materials for ``Inattentive responding can induce spurious associations between task behavior and symptom measures''

### 2023-04-04 manuscripts/supp.tex
checkpointing manuscript progress

% load packages
\usepackage[utf8]{inputenc}
\usepackage[a4paper, margin=0.8in]{geometry}
\usepackage{authblk}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{indentfirst}
\usepackage{xr-hyper} % for cross-referencing between documents
\usepackage{hyperref}
\usepackage{float}
\usepackage{amsmath}
\usepackage[backend=biber,style=nature]{biblatex}
\usepackage[flushmargin]{footmisc}
\usepackage[flushleft]{threeparttable}
\usepackage{array}
\usepackage[font=small, labelfont=it]{caption}
\usepackage{booktabs}
\usepackage{xcolor,soul}
\usepackage[export]{adjustbox}    % center wide tables on page

% specify footnote margins
\renewcommand\footnotelayout{%
  \advance\leftskip 0.7cm
  \advance\rightskip 0.7cm
 }

% setup references
\defbibheading{supp}{}
\addbibresource[label=supp]{supp.bib}
\renewcommand*{\bibfont}{\small}

@@

@@

@@

\maketitle
\begin{refsection}[supp]

@@

\begin{figure}[h]
    \includegraphics[width=17cm]{../figures/supp_01.png}
    \centering
    \captionsetup{width=0.88\textwidth}
    \caption{Signed Spearman rank correlations between task behavior (y-axis) and symptom measures (x-axis) under different regimes of data screening and participant exclusions. (A) No Screening = no exclusions ($N=386$). (B) Accuracy Only = exclusions based on chance-level performance in the reversal-learning task ($N=352$). (C) Infrequency Only = exclusions based on invalid or improbable responses to infrequency items ($N=301$). (D) Both Types = exclusions based on the previous two measures ($N=283$). Only statistically significant correlations are shown ($p<0.05$ not corrected for multiple comparisons). Black Xs indicate significant correlations ablated under screening. Acronyms: Acc = choice accuracy; Pts = total points earned; WS = win-stay rate; LS = lose-shift rate; Pers = perseveration errors; $\beta$ = inverse temperature; $\eta_p$ = positive learning rate; $\eta_n$ = negative learning rate; $\kappa$ = learning rate asymmetry.}
    \label{fig:correlations_s1}
\end{figure}

\clearpage
\subsection*{Bootstrapping Analysis}
\begin{figure}[H]
    \includegraphics[width=16cm]{../figures/supp_02.png}
    \centering
    \captionsetup{width=0.88\textwidth}
    \caption{The pattern of significant behavior-symptom correlations before (A; $N=386$) and after (B; $N=301$) screening compared to the resulting pattern when random subsets of participants ($N=301$, matched to screening using the infrequency measure) are removed (C). Panels (A) and (B) are reproduced from Figure 4 for convenience. (C) The fraction of significant correlations in 5000 bootstrapped samples. (D) The similarity of the pattern of correlations after removal of random subsets to that before and after screening using the infrequency measure ($N=5000$ bootstrapped samples). Similarity was calculated using the simple matching coefficient. Random removal subsets were significantly more similar to the ``No Screening" than to the ``Infrequency" screening datasets (two-tailed, paired-samples $t$-test: $t = 262.490$, $p < 0.001$, $d = 3.713$, 95\% CI = [0.136, 0.138]).}
    \label{fig:random}
\end{figure}

\clearpage
\subsection*{The relationship between sample size and false positive rates generalize to other sets of variables}
\begin{figure}[H]
    \includegraphics[width=16cm]{../figures/supp_03.png}
    \centering
    \captionsetup{width=0.88\textwidth}
    \caption{False positive rates for spurious correlations \emph{increase} with sample size. \textit{Left:} Spearman rank correlations and 95\% bootstrap confidence intervals between inverse temperature ($\beta$) and depression scores (7-down) as a function of sample size and proportion of C/IE participants. The thick dashed lines indicate the threshold for statistical significance for the Spearman correlation at the corresponding sample size. Markers are jittered along the x-axis for legibility. \textit{Right:} False positive rates for inverse temperature ($\beta$) and depression scores (7-down) as a function of sample size and proportion of C/IE participants. False positive rate was calculated as the proportion of bootstrap samples in which the correlation between $\beta$ and 7-down was statistically significant. The horizontal dotted line denotes the expected false positive rate at $\alpha = 0.05$.}
    \label{fig:bootstrap_s1}
\end{figure}

\clearpage
\subsection*{The relationship between sample size and true positive rates for true correlations}
\begin{figure}[H]
    \includegraphics[width=16cm]{../figures/supp_04.png}
    \centering
    \captionsetup{width=0.88\textwidth}
    \caption{True correlations are independent of the proportion of C/IE participants in the sample. \textit{Left:} Spearman rank correlations and 95\% bootstrap confidence intervals between learning rate asymmetry ($\kappa$) and anxiety scores (GAD-7) as a function of sample size and proportion of C/IE participants. The thick dashed lines indicate the threshold for statistical significance for the Spearman correlation at the corresponding sample size. Markers are jittered along the x-axis for legibility. \textit{Right:} True positive rates for learning rate asymmetry ($\kappa$) and anxiety scores (GAD-7) as a function of sample size and proportion of C/IE participants. True positive rate was calculated as the proportion of bootstrap samples in which the correlation between $\kappa$ and GAD-7 was statistically significant. The horizontal dotted line denotes the expected false positive rate at $\alpha = 0.05$.}
    \label{fig:bootstrap_s2}
\end{figure}

@@

\subsection*{Posterior predictive checks}
\begin{figure}[H]
    \includegraphics[width=16cm]{../figures/supp_05.png}
    \centering
    \captionsetup{width=0.88\textwidth}
    \caption{Posterior predictive checks for the risk-sensitive temporal difference learning model. (A) Observed (black) and predicted (blue) learning curves averaged across the group ($N=375$). (B) Observed versus predicted choice accuracy across participants ($N=375$). (C) Distribution of average predictive density across participants ($N=375$).}
    \label{fig:posterior_check}
\end{figure}

\clearpage
\subsection*{Participant demographics}
\begin{table}[!h]
    \centering
    \small
    \setlength{\tabcolsep}{15pt}
    \begin{tabular}{ rrrrr }
        \toprule
        & \multicolumn{2}{c}{MTurk} & \multicolumn{2}{c}{Prolific} \\
        \cmidrule(lr){2-3} \cmidrule(lr){4-5}
        Total & \multicolumn{2}{c}{N=186} & \multicolumn{2}{c}{N=200} \\
        \midrule
        Age & N & \% & N & \% \\
        \cmidrule(lr){1-5}
        18-25   &      16 &   8.6 &       78 &  39.0 \\
        26-35   &      76 &  40.9 &       69 &  34.5 \\
        36-45   &      46 &  24.7 &       31 &  15.5 \\
        46-55   &      22 &  11.8 &       13 &   6.5 \\
        55+     &      26 &  14.0 &        9 &   4.5 \\
        \midrule
        Gender & N & \% & N & \% \\
        \cmidrule(lr){1-5}
        Female             &    83 &  44.6 &    112 &  56.0 \\
        Male               &   103 &  55.4 &     85 &  42.5 \\
        Other              &     0 &   0.0 &      3 &   1.5 \\
        \midrule
        Ethnicity & N & \% & N & \% \\
        \cmidrule(lr){1-5}
        Hispanic or Latino     &    15 &   8.1 &     10 &   5.0 \\
        Not Hispanic or Latino &   168 &  90.3 &    183 &  91.5 \\
        Rather not say         &     2 &   1.1 &      7 &   3.5 \\
        Unknown                &     1 &   0.5 &      0 &   0.0 \\
        \midrule
        Race & N & \% & N & \% \\
        \cmidrule(lr){1-5}
        African American                 &    21 &  11.3 &      7 &   3.5 \\
        Asian                                     &     5 &   2.7 &     53 &  26.5 \\
        White                                     &   151 &  81.2 &    121 &  60.5 \\
        Multiracial                               &     6 &   3.2 &      4 &   2.0 \\
        Rather not say                            &     1 &   0.5 &     12 &   6.0 \\
        \midrule
        Use other platform & N & \% & N & \% \\
        \cmidrule(lr){1-5}
        Yes            &      71 &  38.2 &       28 &  14.0 \\
        No             &     115 &  61.8 &      172 &  86.0 \\
        \bottomrule
    \end{tabular}
    \captionsetup{width=0.88\textwidth}
    \caption{The demographics of each sample by online labor market. On average, the samples were similar though the sample from Mechanical Turk was older (two-tailed, two-sample $t$-test: $t(384)=6.567$, $p<0.001$, $d=0.669$, 95\% CI = [5.4, 10.0]) and comprised of fewer women (two-tailed, two-sample proportions test: $z(384)=2.529$, $p=0.011$, $h=0.258$, 95\% CI = [0.030, 0.228]). Note: the demographics do not include 20 participants excluded for participating in the study twice, once per platform.}
    \label{tab:demographics}
\end{table}

\clearpage
\subsection*{Careless participants show different behaviors on the reversal learning task}
\begin{table}[H]
    \centering
    \begin{tabular}{lcccccc}
    \toprule
    {} &  Attentive &    C/IE &   $t$-value &   $p$-value &  Cohen's $D$ & 95\% CI \\
    \midrule
    Acc      &      0.587 &   0.532 &  4.008 &  $<$0.001 &  0.492 &    [0.028, 0.082] \\
    Pts      &     50.163 &  47.729 &  2.376 &     0.019 &  0.292 &    [0.426, 4.441] \\
    WS       &      0.898 &   0.776 &  5.387 &  $<$0.001 &  0.662 &    [0.077, 0.165] \\
    LS       &      0.609 &   0.751 & -5.335 &  $<$0.001 &  0.655 &  [-0.195, -0.090] \\
    Pers     &      0.245 &   0.259 & -1.505 &     0.139 &  0.185 &   [-0.033, 0.004] \\
    $\beta$  &      6.754 &   4.082 &  5.404 &  $<$0.001 &  0.664 &    [1.702, 3.640] \\
    $\eta_p$ &      0.643 &   0.551 &  2.846 &     0.007 &  0.350 &    [0.029, 0.157] \\
    $\eta_n$ &      0.738 &   0.784 & -1.516 &     0.120 &  0.186 &   [-0.106, 0.013] \\
    $\kappa$ &     -0.069 &  -0.218 &  3.729 &  $<$0.001 &  0.458 &    [0.071, 0.227] \\
    \bottomrule
    \end{tabular}
    \captionsetup{width=0.88\textwidth}
    \caption{Measures of task behavior compared between attentive ($N=301$) and C/IE ($N=85$) participants. Metrics compared using two-tailed, two-sample permutation $t$-test with $df=384$ ($\alpha = 0.05$, not corrected for multiple comparisons). Acronyms: Acc = choice accuracy; Pts = total points earned; WS = win-stay rate; LS = lose-shift rate; Pers = perseveration errors; $\beta$ = inverse temperature; $\eta_p$ = positive learning rate; $\eta_n$ = negative learning rate; $\kappa$ = learning rate asymmetry.}
\end{table}

@@

\subsection*{C/IE responding manifests as qualitatively distinct behavioral strategy}

@@

\begin{figure}[!t]
    \includegraphics[width=16cm]{../figures/main_03c.png}
    \centering
    \captionsetup{width=0.88\textwidth}
    \caption{Softmax regression decision weights and 95\% highest density intervals in attentive (red) and C/IE (blue) participants. The weights reflect the extent to which the recent history of rewards, nonrewards, or previous choices influence current choice. *Denotes where the 95\% highest density interval of the difference in weights excluded zero.}
    \label{fig:regression}
\end{figure}

@@

Participants were recruited as part of two independent studies to investigate reward processing in individuals with and without a history of major depression and bipolar disorder (results of those studies not reported here). Both studies required participants to complete (1) a structured clinical interview (the SCID-5) with a trained interviewer to verify that they met the criteria for one or more psychiatric disorders, and (2) a series of computerized self-report surveys and cognitive tasks.

In both studies, participants volunteered for a multi-session study conducted online via video conferencing. Specifically, participants completed each session from their homes while on Zoom with a study coordinator. During each session, while participants were completing self-report questionnaires or behavioral tasks, the study coordinator turned off their camera and microphone, but was available if the participant had any questions.

@@

51 unique participants were recruited to participate in a behavioral experiment. Of these, 14 participated in both studies. In total, there was data from 20 sessions involving a healthy participant and data from 45 sessions involving a psychiatric participant. Of the 17 healthy participants, 6 identified as men and 11 identified as women. Of the 34 psychiatric patients, 6 identified as men, 27 identified as women, and 1 identified as non-binary. The healthy participants were 24.2 years old on average (sd = 3.5), whereas the psychiatric patients were 27.4 years old on average (sd = 10.6).

@@

\section*{Supplementary references}

\printbibliography[heading=supp]
\end{refsection}

>>>>

% load packages
\usepackage[utf8]{inputenc}
\usepackage[a4paper, margin=0.8in]{geometry}
\usepackage{authblk}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{indentfirst}
\usepackage{hyperref}
\usepackage{float}
\usepackage{amsmath}
\usepackage[backend=biber,style=nature]{biblatex}
\usepackage[flushmargin]{footmisc}
\usepackage[flushleft]{threeparttable}
\usepackage{array}
\usepackage[font=small, labelfont=it]{caption}
\usepackage{booktabs}
\usepackage{xcolor,soul}
\usepackage{adjustbox}

@@

% ----------------------------------%
% Define references
% ----------------------------------%

% setup references
\renewcommand*{\bibfont}{\small}

% define references
\begin{filecontents}{supp.bib}
@article{seymour2012serotonin,
  title={Serotonin selectively modulates reward value in human decision-making},
  author={Seymour, Ben and Daw, Nathaniel D and Roiser, Jonathan P and Dayan, Peter and Dolan, Ray},
  journal={Journal of Neuroscience},
  volume={32},
  number={17},
  pages={5833--5842},
  year={2012},
  publisher={Soc Neuroscience}
}

@ARTICLE{Katahira2018-ke,
  title    = "The statistical structures of reinforcement learning with
              asymmetric value updates",
  author   = "Katahira, Kentaro",
  journal  = "J. Math. Psychol.",
  volume   =  87,
  pages    = "31--45",
  month    =  dec,
  year     =  2018,
  keywords = "Reinforcement learning; Asymmetric value update; Learning rate;
              Choice perseverance; Model fitting; Logistic regression"
}

@article{sugawara2021dissociation,
  title={Dissociation between asymmetric value updating and perseverance in human reinforcement learning},
  author={Sugawara, Michiyo and Katahira, Kentaro},
  journal={Scientific reports},
  volume={11},
  number={1},
  pages={1--13},
  year={2021},
  publisher={Nature Publishing Group}
}

@online{cloudresearch_2020-2,
    author = "Leib Litman",
    title = "New Solutions Dramatically Improve Research Data Quality on {MTurk}",
    url  = "https://www.cloudresearch.com/resources/blog/new-tools-improve-research-data-quality-mturk/",
    addendum = "(Accessed: 2021-02-23)",
    month = jul,
    year = 2020
}

@article{hauser2021evaluating,
  title={Evaluating CloudResearch’s Approved Group as a Solution for Problematic Data Quality on MTurk},
  author={Hauser, David and Moss, Aaron J and Rosenzweig, Cheskie and Jaffe, Shalom Noach and Robinson, Jonathan and LItman, Leib},
  year={2021},
  publisher={PsyArXiv}
}

@MISC{Letzter2021-bw,
  title        = "A teenager on {TikTok} disrupted thousands of scientific
                  studies with a single video",
  author       = "Letzter, Rafi",
  month        =  sep,
  year         =  2021,
  howpublished = "\url{https://www.theverge.com/2021/9/24/22688278/tiktok-science-study-survey-prolific}",
  note         = "Accessed: 2022-10-17"
}

@MISC{prolific_2021,
  title        = "We recently went viral on {TikTok} - here's what we learned",
  howpublished = "\url{https://www.prolific.co/blog/we-recently-went-viral-on-tiktok-heres-what-we-learned}",
  note         = "Accessed: 2022-10-17",
  language     = "en"
}

@article{daw2011model,
  title={Model-based influences on humans' choices and striatal prediction errors},
  author={Daw, Nathaniel D and Gershman, Samuel J and Seymour, Ben and Dayan, Peter and Dolan, Raymond J},
  journal={Neuron},
  volume={69},
  number={6},
  pages={1204--1215},
  year={2011},
  publisher={Elsevier}
}

@article{litman2017turkprime-2,
  title={TurkPrime. com: A versatile crowdsourcing data acquisition platform for the behavioral sciences},
  author={Litman, Leib and Robinson, Jonathan and Abberbock, Tzvi},
  journal={Behavior research methods},
  volume={49},
  number={2},
  pages={433--442},
  year={2017},
  publisher={Springer}
}

@article{robinson2019tapped-2,
  title={Tapped out or barely tapped? Recommendations for how to harness the vast and largely unused potential of the Mechanical Turk participant pool},
  author={Robinson, Jonathan and Rosenzweig, Cheskie and Moss, Aaron J and Litman, Leib},
  journal={PloS one},
  volume={14},
  number={12},
  pages={e0226394},
  year={2019},
  publisher={Public Library of Science San Francisco, CA USA}
}

@article{decker2016creatures,
  title={From creatures of habit to goal-directed learners: Tracking the developmental emergence of model-based reinforcement learning},
  author={Decker, Johannes H and Otto, A Ross and Daw, Nathaniel D and Hartley, Catherine A},
  journal={Psychological science},
  volume={27},
  number={6},
  pages={848--858},
  year={2016},
  publisher={Sage Publications Sage CA: Los Angeles, CA}
}

@article{watson2022development,
  title={The development of preliminary HiTOP internalizing spectrum scales},
  author={Watson, David and Forbes, Miriam K and Levin-Aspenson, Holly F and Ruggero, Camilo J and Kotelnikova, Yuliya and Khoo, Shereen and Bagby, R Michael and Sunderland, Matthew and Patalay, Praveetha and Kotov, Roman},
  journal={Assessment},
  volume={29},
  number={1},
  pages={17--33},
  year={2022},
  publisher={Sage Publications Sage CA: Los Angeles, CA}
}

@article{ashton2007empirical,
  title={Empirical, theoretical, and practical advantages of the HEXACO model of personality structure},
  author={Ashton, Michael C and Lee, Kibeom},
  journal={Personality and social psychology review},
  volume={11},
  number={2},
  pages={150--166},
  year={2007},
  publisher={SAGE publications Sage CA: Los Angeles, CA}
}

@article{deleeuw_2015_jspsych-2,
  title={{jsPsych: A JavaScript} library for creating behavioral experiments in a web browser. },
  author={de Leeuw, J R},
  journal={Behavior Research Methods},
  volume={47},
  number={1},
  pages={1--12},
  year={2015}
}

@article{youngstrom_2013_susd-2,
  title={The {7 Up 7 Down Inventory}: A 14-item measure of manic and depressive tendencies carved from the {General Behavior Inventory}},
  author={Youngstrom, Eric A and Murray, Greg and Johnson, Sheri L and Findling, Robert L},
  journal={Psychological Assessment},
  volume={25},
  number={4},
  pages={1377--1383},
  year={2013},
  publisher={American Psychological Association}
}

@article{winkler2014permutation-2,
  title={Permutation inference for the general linear model},
  author={Winkler, Anderson M and Ridgway, Gerard R and Webster, Matthew A and Smith, Stephen M and Nichols, Thomas E},
  journal={Neuroimage},
  volume={92},
  pages={381--397},
  year={2014},
  publisher={Elsevier}
}

@article{wilcox2018guide-2,
  title={A guide to robust statistical methods in neuroscience},
  author={Wilcox, Rand R and Rousselet, Guillaume A},
  journal={Current protocols in neuroscience},
  volume={82},
  number={1},
  pages={8--42},
  year={2018},
  publisher={Wiley Online Library}
}

@article{harms2018making,
  title={Making'null effects' informative: statistical techniques and inferential frameworks},
  author={Harms, Christopher and Lakens, Dani{\"e}l},
  journal={Journal of Clinical and Translational Research},
  volume={3},
  number={Suppl 2},
  pages={382},
  year={2018},
  publisher={Whioce Publishing Pte. Ltd.}
}

@book{cohen2013statistical,
  title={Statistical power analysis for the behavioral sciences},
  author={Cohen, Jacob},
  year={2013},
  publisher={Routledge}
}

@article{gillan2016characterizing,
  title={Characterizing a psychiatric symptom dimension related to deficits in goal-directed control},
  author={Gillan, Claire M and Kosinski, Michal and Whelan, Robert and Phelps, Elizabeth A and Daw, Nathaniel D},
  journal={elife},
  volume={5},
  pages={e11305},
  year={2016},
  publisher={eLife Sciences Publications Limited}
}

@inproceedings{seabold2010statsmodels,
  title={Statsmodels: Econometric and statistical modeling with python},
  author={Seabold, Skipper and Perktold, Josef},
  booktitle={Proceedings of the 9th Python in Science Conference},
  volume={57},
  number={61},
  pages={10--25080},
  year={2010},
  organization={Austin, TX}
}

@article{barends2019noncompliant-2,
  title={Noncompliant responding: Comparing exclusion criteria in MTurk personality research to improve data quality},
  author={Barends, Ard J and de Vries, Reinout E},
  journal={Personality and individual differences},
  volume={143},
  pages={84--89},
  year={2019},
  publisher={Elsevier}
}

@article{hauser2016attentive-2,
  title={Attentive Turkers: MTurk participants perform better on online attention checks than do subject pool participants},
  author={Hauser, David J and Schwarz, Norbert},
  journal={Behavior research methods},
  volume={48},
  number={1},
  pages={400--407},
  year={2016},
  publisher={Springer}
}

@article{thomas2017validity-2,
  title={Validity and Mechanical Turk: An assessment of exclusion methods and interactive experiments},
  author={Thomas, Kyle A and Clifford, Scott},
  journal={Computers in Human Behavior},
  volume={77},
  pages={184--197},
  year={2017},
  publisher={Elsevier}
}

@article{barch2015mechanisms-2,
  title={Mechanisms underlying motivational deficits in psychopathology: similarities and differences in depression and schizophrenia},
  author={Barch, Deanna M and Pagliaccio, David and Luking, Katherine},
  journal={Behavioral neuroscience of motivation},
  pages={411--449},
  year={2015},
  publisher={Springer}
}

@article{elwert2014endogenous-2,
  title={Endogenous selection bias: The problem of conditioning on a collider variable},
  author={Elwert, Felix and Winship, Christopher},
  journal={Annual review of sociology},
  volume={40},
  pages={31--53},
  year={2014},
  publisher={Annual Reviews}
}

@article{cohen2001impairments-2,
  title={Impairments of attention and effort among patients with major affective disorders},
  author={Cohen, Ronald and Lohr, Ilan and Paul, Robert and Boland, Robert},
  journal={The Journal of neuropsychiatry and clinical neurosciences},
  volume={13},
  number={3},
  pages={385--395},
  year={2001},
  publisher={Am Neuropsych Assoc}
}

@article{culbreth2016negative-2,
  title={Negative symptoms are associated with an increased subjective cost of cognitive effort.},
  author={Culbreth, Adam and Westbrook, Andrew and Barch, Deanna},
  journal={Journal of abnormal psychology},
  volume={125},
  number={4},
  pages={528},
  year={2016},
  publisher={American Psychological Association}
}

@article{kane2016individual-2,
  title={Individual differences in the executive control of attention, memory, and thought, and their associations with schizotypy.},
  author={Kane, Michael J and Meier, Matt E and Smeekens, Bridget A and Gross, Georgina M and Chun, Charlotte A and Silvia, Paul J and Kwapil, Thomas R},
  journal={Journal of Experimental Psychology: General},
  volume={145},
  number={8},
  pages={1017},
  year={2016},
  publisher={American Psychological Association}
}

@article{robison2017neurotic-2,
  title={The neurotic wandering mind: An individual differences investigation of neuroticism, mind-wandering, and executive control},
  author={Robison, Matthew K and Gath, Katherine I and Unsworth, Nash},
  journal={The Quarterly Journal of Experimental Psychology},
  volume={70},
  number={4},
  pages={649--663},
  year={2017},
  publisher={Taylor \& Francis}
}
\end{filecontents}
\addbibresource{supp.bib}

% ----------------------------------%
% Define author / abstract block
% ----------------------------------%

@@

% ----------------------------------%
% Main text
% ----------------------------------%

\maketitle

@@

@@

\begin{figure}[H]
    \includegraphics[width=17cm]{../figures/supp_01.png}
    \centering
    \captionsetup{width=0.88\textwidth}
    \caption{Signed Spearman rank correlations between task behavior (y-axis) and symptom measures (x-axis) under different regimes of data screening and participant exclusions. (A) No Screening = no exclusions ($N=386$). (B) Accuracy Only = exclusions based on chance-level performance in the reversal-learning task ($N=352$). (C) Infrequency Only = exclusions based on invalid or improbable responses to infrequency items ($N=301$). (D) Both Types = exclusions based on the previous two measures ($N=283$). Only statistically significant correlations are shown ($p<0.05$ not corrected for multiple comparisons). Black Xs indicate significant correlations ablated under screening. Acronyms: Acc = choice accuracy; Pts = total points earned; WS = win-stay rate; LS = lose-shift rate; Pers = perseveration errors; $\beta$ = inverse temperature; $\eta_p$ = positive learning rate; $\eta_n$ = negative learning rate; $\kappa$ = learning rate asymmetry.}
    \label{fig:figS01}
\end{figure}

\clearpage
\subsection*{Bootstrapping Analysis}
\begin{figure}[H]
    \includegraphics[width=16cm]{../figures/supp_02.png}
    \centering
    \captionsetup{width=0.88\textwidth}
    \caption{The pattern of significant behavior-symptom correlations before (A; $N=386$) and after (B; $N=301$) screening compared to the resulting pattern when random subsets of participants ($N=301$, matched to screening using the infrequency measure) are removed (C). Panels (A) and (B) are reproduced from Figure 4 for convenience. (C) The fraction of significant correlations in 5000 bootstrapped samples. (D) The similarity of the pattern of correlations after removal of random subsets to that before and after screening using the infrequency measure ($N=5000$ bootstrapped samples). Similarity was calculated using the simple matching coefficient. Random removal subsets were significantly more similar to the ``No Screening" than to the ``Infrequency" screening datasets (two-tailed, paired-samples $t$-test: $t = 262.490$, $p < 0.001$, $d = 3.713$, 95\% CI = [0.136, 0.138]).}
    \label{fig:figS02}
\end{figure}

\clearpage
\subsection*{The relationship between sample size and false positive rates generalize to other sets of variables}
\begin{figure}[H]
    \includegraphics[width=16cm]{../figures/supp_03.png}
    \centering
    \captionsetup{width=0.88\textwidth}
    \caption{False positive rates for spurious correlations \emph{increase} with sample size. \textit{Left:} Spearman rank correlations and 95\% bootstrap confidence intervals between inverse temperature ($\beta$) and depression scores (7-down) as a function of sample size and proportion of C/IE participants. The thick dashed lines indicate the threshold for statistical significance for the Spearman correlation at the corresponding sample size. Markers are jittered along the x-axis for legibility. \textit{Right:} False positive rates for inverse temperature ($\beta$) and depression scores (7-down) as a function of sample size and proportion of C/IE participants. False positive rate was calculated as the proportion of bootstrap samples in which the correlation between $\beta$ and 7-down was statistically significant. The horizontal dotted line denotes the expected false positive rate at $\alpha = 0.05$.}
    \label{fig:figS03}
\end{figure}

\clearpage
\subsection*{The relationship between sample size and true positive rates for true correlations}
\begin{figure}[H]
    \includegraphics[width=16cm]{../figures/supp_04.png}
    \centering
    \captionsetup{width=0.88\textwidth}
    \caption{True correlations are independent of the proportion of C/IE participants in the sample. \textit{Left:} Spearman rank correlations and 95\% bootstrap confidence intervals between learning rate asymmetry ($\kappa$) and anxiety scores (GAD-7) as a function of sample size and proportion of C/IE participants. The thick dashed lines indicate the threshold for statistical significance for the Spearman correlation at the corresponding sample size. Markers are jittered along the x-axis for legibility. \textit{Right:} True positive rates for learning rate asymmetry ($\kappa$) and anxiety scores (GAD-7) as a function of sample size and proportion of C/IE participants. True positive rate was calculated as the proportion of bootstrap samples in which the correlation between $\kappa$ and GAD-7 was statistically significant. The horizontal dotted line denotes the expected false positive rate at $\alpha = 0.05$.}
    \label{fig:figS04}
\end{figure}

@@

\subsection*{Posterior predictive checks}
\begin{figure}[H]
    \includegraphics[width=16cm]{../figures/supp_05.png}
    \centering
    \captionsetup{width=0.88\textwidth}
    \caption{Posterior predictive checks for the risk-sensitive temporal difference learning model. (A) Observed (black) and predicted (blue) learning curves averaged across the group ($N=375$). (B) Observed versus predicted choice accuracy across participants ($N=375$). (C) Distribution of average predictive density across participants ($N=375$).}
    \label{fig:figS05}
\end{figure}

\clearpage
\subsection*{Participant demographics}
\begin{table}[!h]
    \centering
    \small
    \setlength{\tabcolsep}{15pt}
    \begin{tabular}{ rrrrr }
        \toprule
        & \multicolumn{2}{c}{MTurk} & \multicolumn{2}{c}{Prolific} \\
        \cmidrule(lr){2-3} \cmidrule(lr){4-5}
        Total & \multicolumn{2}{c}{N=186} & \multicolumn{2}{c}{N=200} \\
        \midrule
        Age & N & \% & N & \% \\
        \cmidrule(lr){1-5}
        18-25   &      16 &   8.6 &       78 &  39.0 \\
        26-35   &      76 &  40.9 &       69 &  34.5 \\
        36-45   &      46 &  24.7 &       31 &  15.5 \\
        46-55   &      22 &  11.8 &       13 &   6.5 \\
        55+     &      26 &  14.0 &        9 &   4.5 \\
        \midrule
        Gender & N & \% & N & \% \\
        \cmidrule(lr){1-5}
        Female             &    83 &  44.6 &    112 &  56.0 \\
        Male               &   103 &  55.4 &     85 &  42.5 \\
        Other              &     0 &   0.0 &      3 &   1.5 \\
        \midrule
        Ethnicity & N & \% & N & \% \\
        \cmidrule(lr){1-5}
        Hispanic or Latino     &    15 &   8.1 &     10 &   5.0 \\
        Not Hispanic or Latino &   168 &  90.3 &    183 &  91.5 \\
        Rather not say         &     2 &   1.1 &      7 &   3.5 \\
        Unknown                &     1 &   0.5 &      0 &   0.0 \\
        \midrule
        Race & N & \% & N & \% \\
        \cmidrule(lr){1-5}
        African American                 &    21 &  11.3 &      7 &   3.5 \\
        Asian                                     &     5 &   2.7 &     53 &  26.5 \\
        White                                     &   151 &  81.2 &    121 &  60.5 \\
        Multiracial                               &     6 &   3.2 &      4 &   2.0 \\
        Rather not say                            &     1 &   0.5 &     12 &   6.0 \\
        \midrule
        Use other platform & N & \% & N & \% \\
        \cmidrule(lr){1-5}
        Yes            &      71 &  38.2 &       28 &  14.0 \\
        No             &     115 &  61.8 &      172 &  86.0 \\
        \bottomrule
    \end{tabular}
    \captionsetup{width=0.88\textwidth}
    \caption{The demographics of each sample by online labor market. On average, the samples were similar though the sample from Mechanical Turk was older (two-tailed, two-sample $t$-test: $t(384)=6.567$, $p<0.001$, $d=0.669$, 95\% CI = [5.4, 10.0]) and comprised of fewer women (two-tailed, two-sample proportions test: $z(384)=2.529$, $p=0.011$, $h=0.258$, 95\% CI = [0.030, 0.228]). Note: the demographics do not include 20 participants excluded for participating in the study twice, once per platform.}
    \label{tab:tabS01}
\end{table}

\clearpage
\subsection*{Careless participants show different behaviors on the reversal learning task}
\begin{table}[H]
    \centering
    \begin{tabular}{lcccccc}
    \toprule
    {} &  Attentive &    C/IE &   $t$-value &   $p$-value &  Cohen's $D$ & 95\% CI \\
    \midrule
    Acc      &      0.587 &   0.532 &  4.008 &  $<$0.001 &  0.492 &    [0.028, 0.082] \\
    Pts      &     50.163 &  47.729 &  2.376 &     0.019 &  0.292 &    [0.426, 4.441] \\
    WS       &      0.898 &   0.776 &  5.387 &  $<$0.001 &  0.662 &    [0.077, 0.165] \\
    LS       &      0.609 &   0.751 & -5.335 &  $<$0.001 &  0.655 &  [-0.195, -0.090] \\
    Pers     &      0.245 &   0.259 & -1.505 &     0.139 &  0.185 &   [-0.033, 0.004] \\
    $\beta$  &      6.754 &   4.082 &  5.404 &  $<$0.001 &  0.664 &    [1.702, 3.640] \\
    $\eta_p$ &      0.643 &   0.551 &  2.846 &     0.007 &  0.350 &    [0.029, 0.157] \\
    $\eta_n$ &      0.738 &   0.784 & -1.516 &     0.120 &  0.186 &   [-0.106, 0.013] \\
    $\kappa$ &     -0.069 &  -0.218 &  3.729 &  $<$0.001 &  0.458 &    [0.071, 0.227] \\
    \bottomrule
    \end{tabular}
    \captionsetup{width=0.88\textwidth}
    \caption{Measures of task behavior compared between attentive ($N=301$) and C/IE ($N=85$) participants. Metrics compared using two-tailed, two-sample permutation $t$-test with $df=384$ ($\alpha = 0.05$, not corrected for multiple comparisons). Acronyms: Acc = choice accuracy; Pts = total points earned; WS = win-stay rate; LS = lose-shift rate; Pers = perseveration errors; $\beta$ = inverse temperature; $\eta_p$ = positive learning rate; $\eta_n$ = negative learning rate; $\kappa$ = learning rate asymmetry.}
    \label{tab:tabS02}
\end{table}

@@

\subsection*{C/IE responding manifests as a distinct behavioral strategy}

@@

\begin{figure}[!t]
    \includegraphics[width=16cm]{../figures/main_03c.png}
    \centering
    \caption{Softmax regression decision weights and 95\% highest density intervals in attentive (red) and C/IE (blue) participants. The weights reflect the extent to which the recent history of rewards, nonrewards, or previous choices influence current choice. *Denotes where the 95\% highest density interval of the difference in weights excluded zero.}
    \label{fig:regression}
\end{figure}

@@

Participants were recruited as part of two independent studies to investigate reward processing in individuals with and without a history of major depression and bipolar disorder (results of those studies not reported here). The studies were approved by the Institutional Review Board of Rutgers University (\#2019000738, \#2018000629), and all participants provided informed consent. Participants received monetary compensation for their time (rate USD \$12/hr), plus an incentive-compatible bonus based on task performance.

In both studies, participants volunteered for a multi-session study conducted online via video conferencing. Specifically, participants completed each session from their homes while on Zoom with a study coordinator. Both studies required participants to complete (1) a structured clinical interview (the SCID-5) with a trained interviewer to verify that they met the criteria for one or more psychiatric disorders, and (2) a series of computerized self-report surveys and cognitive tasks. During each session, while participants were completing self-report questionnaires or behavioral tasks, the study coordinator turned off their camera and microphone, but was available if the participant had any questions.

@@

52 unique participants were recruited to participate in a behavioral experiment. Of these, 14 participated in both studies. In total, there was data from 20 sessions involving a healthy participant and data from 45 sessions involving a psychiatric participant. Of the 17 healthy participants, 6 identified as men and 11 identified as women. Of the 35 psychiatric patients, 6 identified as men, 27 identified as women, 1 identified as non-binary, and 1 preferred not to say. The healthy participants were 24.2 years old on average (sd = 3.5), whereas the psychiatric patients were 27.4 years old on average (sd = 10.6).

@@

\printbibliography[title=Supplementary References]
